---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r loaddata,echo=TRUE}
columnClasses <- c("numeric", "character", "numeric")
activityData <- read.csv("activity.csv", colClasses = columnClasses)
```

## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day.
```{r dailydata, echo=TRUE}
dailyActivityData <- aggregate(steps ~ date, activityData, sum)
```
2. Histogram of total steps taken per day.
```{r histogram, echo=TRUE}
par(mfrow = c(1, 1))
hist(dailyActivityData$steps, main = "Total Steps Taken Per Day", xlab = "Daily steps")
```
3. Calculate the mean and median
```{r stats, echo=TRUE}
mean(dailyActivityData$steps, na.rm = TRUE)
median(dailyActivityData$steps, na.rm = TRUE)
```

## What is the average daily activity pattern?

1. Time series plot of the 5-minute interval vs avg number of steps taken per day.

```{r timeseries, echo=TRUE}
stepActivityData <- aggregate(steps ~ interval, activityData, mean)
plot(stepActivityData$interval, stepActivityData$steps, type = "l", main = "Avg steps at each interval", xlab = "Interval", ylab = "Avg steps per day")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

Interval `r stepActivityData[stepActivityData$steps == max(stepActivityData$steps),]$interval` has the most number of average steps, `r max(stepActivityData$steps)`.

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
